---
title: "Master Thesis: Effect of Number of partners on International Joint Ventures Survival: the moderating role of Environmental Dynamism & Technology Transfer"
output:
  html_document:
    df_print: paged
    theme: united
    highlight: textmate
  word_document: default
  pdf_document: default
---

1. Importing necessary libraries

```{r, eval=FALSE}
# Install and load necessary packages
if (!require("tidyverse")) install.packages("tidyverse", dependencies=TRUE)
if (!require("survival")) install.packages("survival", dependencies=TRUE)
if (!require("caret")) install.packages("caret", dependencies=TRUE)
if (!require("ggplot2")) install.packages("ggplot2", dependencies=TRUE)
if (!require("readr")) install.packages("readr", dependencies=TRUE)
if (!require("stargazer")) install.packages("stargazer", dependencies=TRUE)
if (!require("car")) install.packages("car", dependencies=TRUE)
if (!require("psych")) install.packages("psych", dependencies=TRUE)
if (!require("dplyr")) install.packages("dplyr", dependencies=TRUE)
if (!require("tidyr")) install.packages("tidyr", dependencies=TRUE)
if (!require("flextable")) install.packages("flextable", dependencies=TRUE)
if (!require("officer")) install.packages("officer", dependencies=TRUE)
if (!require("survminer")) install.packages("survminer", dependencies=TRUE)
if (!require("htmlwidgets")) install.packages("htmlwidgets", dependencies=TRUE)
if (!require("scales")) install.packages("scales", dependencies=TRUE)

# Load required libraries
library(tidyverse)
library(survival)
library(caret)
library(ggplot2)
library(readr)
library(stargazer)
library(car)
library(psych)
library(dplyr)
library(tidyr)
library(flextable)
library(officer)
library("survminer")
library("htmlwidgets")
library(scales)
```

2. View citation of the survival package used in "R"

```{r, eval=FALSE}
# Citation and package version of survival package
citation("survival")
packageVersion("survival")
```

3. Read & View source file from the directory 

```{r, eval=FALSE}

df <- read_csv("IJV_Sample_Cox_Model.csv") #Replace the source file path name & file name.

# View the DataFrame
print(head(df))

# Drop the "Deal Name" column
df <- df %>% 
  select(-'Deal Name')

# View the DataFrame after dropping the column
print(head(df))
```

4. Prepare the data by converting variables to appropriate formats for analysis

```{r, eval=FALSE}
# Prepare Data for Analysis
df <- df %>%
  mutate(termination = as.numeric(termination),
         number_of_partners = as.numeric(number_of_participants),
         technology_transfer = as.numeric(technology_transfer),
         ijv_scope = as.numeric(ijv_scope),
         industry_relatedness = as.numeric(industry_relatedness),
         start_year = as.factor(start_year),
         industry_sic = as.factor(industry_sic),
         ijv_age = as.numeric(ijv_age),
         partner_countries = as.numeric(partner_countries))
```

5. Normalize numeric variables to improve model performance

```{r, eval=FALSE}

# Normalize numeric variables
numeric_vars <- c("environmental_dynamism", "cultural_distance", "gdp_difference")
preProcValues <- preProcess(df[numeric_vars], method = c("center", "scale"))
df[numeric_vars] <- predict(preProcValues, df[numeric_vars])
```

6. Outlier detection and handling

```{r, eval=FALSE}

# Check for outliers
outliers <- df %>%
  select(numeric_vars) %>%
  gather(key = "variable", value = "value") %>%
  group_by(variable) %>%
  mutate(outlier = ifelse(value < quantile(value, 0.25) - 1.5 * IQR(value) | 
                            value > quantile(value, 0.75) + 1.5 * IQR(value), TRUE, FALSE))

# Plot outliers before handling
options(warn= 0)
ggplot(outliers, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_jitter(aes(color = outlier), width = 0.2) +
  theme_minimal()

# Handle outliers by capping them at the 5th and 95th percentiles
for (var in numeric_vars) {
  lower_bound <- quantile(df[[var]], 0.05)
  upper_bound <- quantile(df[[var]], 0.95)
  df[[var]] <- ifelse(df[[var]] < lower_bound, lower_bound, df[[var]])
  df[[var]] <- ifelse(df[[var]] > upper_bound, upper_bound, df[[var]])
}

# Check for outliers again after handling
outliers_handled <- df %>%
  select(numeric_vars) %>%
  gather(key = "variable", value = "value") %>%
  group_by(variable) %>%
  mutate(outlier = ifelse(value < quantile(value, 0.25) - 1.5 * IQR(value) | 
                            value > quantile(value, 0.75) + 1.5 * IQR(value), TRUE, FALSE))

# Plot outliers after handling
ggplot(outliers_handled, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_jitter(aes(color = outlier), width = 0.2) +
  theme_minimal() +
  labs(title = "Outliers in Numeric Variables (After Handling)", x = "Variable", y = "Value")
```

7. Generate Descriptive Statistics & Correlation Matrix

```{r, eval=FALSE}

# Descriptive stats and correlation matrix
# Define the variables for the correlation matrix
vars_for_corr <- c("termination","ijv_age","number_of_partners", "environmental_dynamism",
                   "technology_transfer", "cultural_distance", "gdp_difference",
                   "ijv_scope", "industry_relatedness", "partner_countries")

# Calculate means and standard deviations and round to 3 digits
means <- round(colMeans(df[vars_for_corr], na.rm = TRUE), 3)
sds <- round(apply(df[vars_for_corr], 2, sd, na.rm = TRUE), 3)

# Calculate correlation matrix with significance levels
correlation_results <- corr.test(df[vars_for_corr], use = "pairwise.complete.obs", method = "pearson")

# Extract the correlation matrix and p-values
cor_matrix <- correlation_results$r
p_values <- correlation_results$p

# Create a data frame for means and standard deviations
summary_stats <- data.frame(Serial = 1:length(vars_for_corr),Variable = vars_for_corr, Mean = means, SD = sds)

# Prepare the correlation matrix with significance levels
stars <- symnum(p_values, cutpoints = c(0, 0.001, 0.01, 0.05, 1), symbols = c("***", "**", "*", ""))
cor_matrix_with_stars <- cor_matrix
for (i in seq_len(ncol(cor_matrix))) {
  for (j in seq_len(nrow(cor_matrix))) {
    cor_matrix_with_stars[i, j] <- paste0(formatC(cor_matrix[i, j], format = "f", digits = 2), " ", stars[i, j])
  }
}

# Convert the matrix to a data frame for better display
cor_matrix_df <- as.data.frame(cor_matrix_with_stars)
colnames(cor_matrix_df) <- vars_for_corr
rownames(cor_matrix_df) <- vars_for_corr

# Combine the summary statistics and correlation matrix
combined_results <- summary_stats %>%
  bind_cols(cor_matrix_df)

# Create a flextable object with header labels
ft <- flextable(combined_results) %>%
  set_header_labels(Serial = "S.No",
                    Variable = "Variable", 
                    Mean = "Mean", 
                    SD = "SD", 
                    termination = "1",
                    ijv_age = "2",
                    number_of_partners = "3",
                    environmental_dynamism = "4",
                    technology_transfer = "5",
                    cultural_distance = "6",
                    gdp_difference = "7",
                    ijv_scope = "8",
                    industry_relatedness = "9",
                    partner_countries = "10") %>%
  add_header_row(values = c("S.No", "Variable", "Mean", "SD", "Correlation Matrix"), colwidths = c(1, 1, 1, 1, length(vars_for_corr))) %>%
  add_footer_lines(values = "Note: *** p < 0.001, ** p < 0.01, * p < 0.05") %>%
  autofit()

# View the flextable in RStudio viewer
ft
```

8. Perform Variation Inflation Factors (VIF) test to check for multicollinearity among covariates

```{r, eval=FALSE}

# Define a formula for VIF calculation with only independent variables, moderators, and controls
formula_vif <- as.formula("termination ~ number_of_partners + technology_transfer + environmental_dynamism + cultural_distance + gdp_difference + ijv_scope + industry_relatedness + partner_countries")
linear_model <- lm(formula_vif, data = df)

# Calculate VIF
vif_results <- vif(linear_model)
vif_results

# Round VIF values to 3 digits
vif_results <- round(vif_results, 3)
vif_results

# Create a data frame for VIF values
vif_table <- data.frame(Variable = names(vif_results), VIF = vif_results)

# Reorder the data frame to have independent variables and moderators at the start
independent_vars <- c("number_of_partners", "technology_transfer", "environmental_dynamism")
control_vars <- c("cultural_distance", "gdp_difference", "ijv_scope", "industry_relatedness","partner
                  _countries")
vif_table <- vif_table %>%
  mutate(Order = match(Variable, c(independent_vars, control_vars))) %>%
  arrange(Order) %>%
  select(-Order)

# Create a flextable object and rename columns
ft_vif <- flextable(vif_table) %>%
  set_header_labels(Variable = "Variable", VIF = "VIF") %>%
  autofit()

# Optional: Print the table to view it in RStudio viewer
ft_vif
```

9. To verify whether Cox Proportional Hazard assumptions hold for covariates, Schoenfeld Global test is conducted.

```{r, eval=TRUE}

# Check for Cox proportional hazards (PH) model assumptions without including fixed effects
# Model for Hypothesis 1 (excluding fixed effects)
model_h1 <- coxph(Surv(time = ijv_age, event = termination) ~ 
                           number_of_participants*environmental_dynamism +
                           cultural_distance + gdp_difference + 
                           ijv_scope + industry_relatedness + partner_countries, 
                         data = df)
# Model for Hypothesis 2 (excluding fixed effects)
model_h2 <- coxph(Surv(time = ijv_age, event = termination) ~ 
                           number_of_participants * technology_transfer +
                           cultural_distance + gdp_difference + 
                           ijv_scope + industry_relatedness + partner_countries, 
                         data = df)

# Check PH assumption for Hypothesis 1
ph_test_h1 <- cox.zph(model_h1)
print(ph_test_h1)
plot(ph_test_h1)
print(names(ph_test_h1))
ggcoxzph(ph_test_h1)

# Plotting graph for EVD covariate (included in Thesis document)
ggcoxzph(ph_test_h1, resid = TRUE, se = TRUE, df = 4, var = 2,font.main = 20, point.size = 4, point.col = "blue",ggtheme =
  theme(
    text = element_text(family = "Times New Roman", size = 20),  # Set font to Times New Roman
    plot.title = element_text(family = "Times New Roman", size = 20),
    axis.title = element_text(family = "Times New Roman", size = 20),
    axis.text = element_text(family = "Times New Roman", size = 20)
  ))

# Check PH assumption for Hypothesis 2
ph_test_h2 <- cox.zph(model_h2)
print(ph_test_h2)
plot(ph_test_h2)
print(names(ph_test_h2))
ggcoxzph(ph_test_h2)
```

10. Perform Cox regression analysis for testing Hypothesis 1 & 2

```{r, eval=FALSE}

# Perform Cox regression with fixed effects
# Model 1
cox_model_h1 <- coxph(Surv(ijv_age, termination) ~ number_of_partners*environmental_dynamism + 
                        cultural_distance + gdp_difference + 
                        ijv_scope + industry_relatedness + partner_countries + factor(industry_sic) + 
                        factor(start_year),
                        data = df)

# Model 2
cox_model_h2 <- coxph(Surv(ijv_age, termination) ~ 
                        number_of_partners*technology_transfer +
                        cultural_distance + gdp_difference + 
                        ijv_scope  + industry_relatedness  + partner_countries + factor(industry_sic) + 
                        factor(start_year),
                      data = df)

# Setting max.print to a high value
options(max.print = 500000)

# Summarize the models
summary_H1 <- summary(cox_model_h1)
summary_H1
summary_H2 <- summary(cox_model_h2)
summary_H2

# Export the results using stargazer, for quick interpretation and structured overview
stargazer(cox_model_h1, cox_model_h2, type = "text",
          title = "Cox Regression Results",
          dep.var.labels = c("Proportional Hazard Model:Termination = 1"),
          covariate.labels = c("Number of Participants", "Environmental Dynamism", "Technology Transfer", 
                               "Cultural Distance", "GDP Difference", "IJV Scope", 
                               "Industry Relatedness", "Partner Countries", "Industry SIC", "Start Year"),
          omit.stat = c("wald", "lr"),
          out = "cox_regression_results.html")

```

11. Plotting survival curves to visualise the interaction effect defined in Hypothesis 1 & 2.

```{r, eval=FALSE}

# Define a custom theme for formatting the visuals
custom_theme <- theme_minimal() +
  theme(
    text = element_text(family = "Times", size = 114),
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 14,face = "bold"),
    legend.text = element_text(size = 12),
    plot.background = element_rect(fill = "white"),
    panel.grid = element_line(color = "grey80", linewidth = 0.1),
    panel.grid.minor = element_blank(),
    legend.position = "bottom" 
  )

# Survival Plot for Hypothesis 1: Interaction between Number of partners & Environemntal dynamism
# Convert to survival probabilities (assuming 1 - mean_event gives survival probability)
interaction_h1 <- df %>% 
  mutate(predicted_event = predict(cox_model_h1, type = "expected")) %>%
  group_by(number_of_participants, environmental_dynamism) %>%
  summarise(mean_event = mean(predicted_event),.groups = 'drop') %>%
  mutate(mean_survival_prob = exp(-mean_event) * 100)  # Convert to percentage

# Plot with improved clarity and color scheme
ggplot(interaction_h1, aes(x = as.numeric(number_of_participants), y = mean_survival_prob, color = environmental_dynamism)) +
  geom_line() +
  scale_color_gradient2(low = "red", mid = "blue", high = "green", midpoint = median(df$environmental_dynamism), space = "Lab", na.value = "grey50")  +
  labs(
    x = "Number of Partners",
    y = "Mean Survival Probability (%)",
    color = "Environmental Dynamism Index") +
  custom_theme

# Survival Plot for Hypothesis 1: Interaction between Number of partners & Technology Transfer
interaction_h2 <- df %>% 
  mutate(predicted_event = predict(cox_model_h2, type = "expected")) %>%
  group_by(number_of_participants, technology_transfer) %>%
  summarise(mean_event = mean(predicted_event),.groups = 'drop')

# Convert to survival probabilities (assuming 1 - mean_event gives survival probability)
interaction_h2 <- interaction_h2 %>% 
  #mutate(mean_survival_prob = (1 - mean_event) * 100)  # Convert to percentage
  mutate(mean_survival_prob = exp(-mean_event) * 100)

# Plot with improved clarity and color scheme
ggplot(interaction_h2, aes(x = as.numeric(number_of_participants), y = mean_survival_prob, color = factor(technology_transfer))) +
  geom_line(linewidth = 0.5) +  # Use linewidth instead of size
  geom_point(size = 1) +
  scale_color_manual(values = c("1" = "green", "0" = "blue")) +
  labs(
    x = "Number of Partners",
    y = "Mean Survival Probability (%)",
    color = "Technology Transfer" ) +
  custom_theme
```


                      






